STA302 Fall 2024 Final Project Part 1: Research Proposal and Data Introduction

### 2024-10-07 final_paper/paper.qmd
Polished and prepared to submit

# Preliminary Results 
Our model demonstrates the relationship between the New Housing Price Index (NHPI) and various predictors, such as detached absorption and unabsorbed homes, homes starting construction, completed constructions, and rate hikes. To start, we will discuss the intercept, which is 88.52 when all predictors are zero. This serves as a baseline for evaluating the effects of other predictors.

The first predictor is detached absorption. For each additional home sold in a quarter, the expected decrease in the NHPI is 0.0227 points. In contrast, unabsorbed homes cause an increase of 0.005315 points in the NHPI. New homes beginning construction have a minimal impact, with a 0.0009611 point increase in the NHPI for every home that starts construction. Meanwhile, completed homes have a greater impact, with a 0.002977 point increase. Lastly, rate hikes have the most significant impact: when there is a rate increase, the NHPI decreases by 13.41 points.

## Residual Analysis
In this section, we will explore the assumptions made for our linear regression models and check for any violations. To begin, using our model in @sec-model, we created residual plots to verify these assumptions. The first graph we will examine is @fig-resid, which is our Residuals vs. Fitted Values plot. This graph shows some clustering around a fitted value of 100 and above the 0 residual line. It also displays a decreasing fanning pattern. These two characteristics suggest potential non-linearity and some correlated errors.

Next, we will review the residuals vs. predictors. In @fig-residuals-3 and @fig-residuals-4 The plots for homes being built and completed homes show the least amount of assumption violations, as the graphs exhibit good spread and randomness. In contrast, the plots for homes sold @fig-residuals-1 and unsold homes @fig-residuals-2 display some fanning and clustering, which may indicate correlated errors and non-constant variance. Our categorical predictor, rate hike @fig-residuals-5 , reveals some non-constant variance, as the spread of residuals for "1" is wider than for "0."

Lastly, while most residuals align well with the line in the QQ plot @fig-qq_plot , there are deviations at the lower and upper extremes, suggesting some skewness. This indicates that the residuals are not perfectly normally distributed, which may signal a violation of the normality assumption.

>>>>

# Preliminary Results 
Our model examines the relationship between the New Housing Price Index (NHPI) and predictors like detached absorption, unabsorbed homes, new homes under construction, completed constructions, and rate hikes. The intercept is 88.52, representing the NHPI value when all predictors are zero, serving as a baseline for evaluating the effects of other variables.

Detached absorption, our first predictor, shows that each additional home sold in a quarter is associated with a decrease of 0.0227 points in the NHPI. Conversely, unabsorbed homes increase the NHPI by 0.005315 points. The impact of new homes under construction is minimal, increasing the NHPI by 0.0009611 points for each home started. Completed constructions have a greater effect, contributing a 0.002977 point increase. Rate hikes, however, have the most significant influence: each rate increase leads to a 13.41-point decrease in the NHPI.

## Residual Analysis
We assessed the assumptions of our linear regression model by examining residual plots. The Residuals vs. Fitted Values plot (@fig-resid) reveals clustering around a fitted value of 100, with a slight decrease in spread, indicating potential non-linearity and correlated errors.

Further, we reviewed residuals vs. predictors. The residual plots for homes under construction and completed homes (@fig-residuals-3 and @fig-residuals-4) show minimal violations, as the spread appears random and well-distributed. However, the residuals for sold (@fig-residuals-1) and unsold homes (@fig-residuals-2) display fanning and clustering, suggesting correlated errors and non-constant variance. Additionally, the categorical predictor "rate hike" (@fig-residuals-5) shows wider residual variance for a value of "1" than for "0," indicating non-constant variance.

Lastly, while the QQ plot (@fig-qq_plot) suggests that most residuals align with the expected normal distribution, deviations at the lower and upper extremes indicate some skewness. This suggests that the residuals are not perfectly normally distributed, hinting at a potential violation of the normality assumption.
